 Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation

### 2021-06-09 analysis/paper/shceof.Rmd
Arregla figura extended-series
#46

panel_background <- theme(panel.background = element_rect(fill = "#fbfbfb", color = NA), 
                          panel.ontop = FALSE,
                          tagger.panel.tag.background = ggplot2::element_rect(color = NA, 
                                                                              fill = "#fbfbfb"))

@@

```{r ceofs-2, fig.cap = "Time series of the leading two cEOFs of zonal anomalies of geopotential height at 50 hPa and 200 hPa for the SON trimester and the period 1979 -- 2019.", fig.height=width_column, fig.width = width_column}
ceof %>%
  .[, eof[[1]]$left, by = season] %>%
  sep_ReIm(hgt) %>%
  .[season == "SON"] %>%
  # .[, hgt := scale(hgt), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt)) +
  geom_hline(yintercept = 0, size = 0.1, color = "gray30") +
  geom_line(aes(color = part)) +
  scale_x_datetime(NULL) +
  scale_y_continuous(NULL) +
  scale_color_brewer(NULL, palette = "Dark2") +
  facet_grid(cEOF ~ ., switch = "y") +
  theme(panel.spacing = grid::unit(1.5, "lines"), strip.placement = "outside")  +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  panel_background
```

@@

(ref:extended-series-cap) Time series extended using ERA5 back extended preliminary edition (period 1950 -- 1978) and ERA5 (period 1979 -- 2019).
Each series is scaled to zero mean and unit standard deviation.

```{r extended-series, fig.cap = "(ref:extended-series-cap)", fig.height=3, fig.width = 7, fig.env = "figure*"}
# Check that the extended index coincides with the
# original index.
cors <- era_more %>%
  .[sep_ReIm(ceof$eof[[1]]$left), on = c("time", "cEOF", "part")] %>%
  na.omit() %>%
  .[, cor(hgt, i.hgt), by = .(cEOF, part)]

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
   # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.1, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  geom_line(aes(color = part)) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = part)) +
  scale_x_datetime(NULL, date_labels = "%Y") +
  scale_y_continuous("EOF") +
  scale_color_brewer(NULL, palette = "Dark2") +
  facet_grid(cEOF ~ part, scales = "free_y") +
  theme(panel.spacing = grid::unit(1.5, "lines"))  +
  tag_facets("cr") +
  panel_background
```

@@

\textcolor{red}{Revisa la figura 4 que te están quedando bastante desparejos los subplots. También sugieron mantener la convención de colores de la figura anterior (real/imaginario) para hacer más directa la interpretación.}

@@

The spatial patterns shown in Figure \@ref(fig:ceofs-1) are derived by removing the zonally symmetric circulation, so they might not include all the variability that is actually associated with the cEOF time series.
To understand the full geopotential height field changes associated with each cEOF, we computed regression patterns of each cEOF with the complete geopotential fields.

@@

  scale_x_level() +
  scale_y_continuous(r2, limits = c(0, 1)) +
  scale_size_manual("p-value", values = c("TRUE" = 1, 
                                      "FALSE" = 0.5), 
                    guide = "none") +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ cEOF) +
  coord_flip() +
  tag_facets("rc") +
  panel_background
```

>>>>

panel_background <- theme(panel.background = element_rect(fill = "#fbfbfb", color = NA), 
                          panel.ontop = FALSE,
                          panel.grid = element_line(size = 0.5, color = "gray90"),
                          tagger.panel.tag.background = ggplot2::element_rect(color = NA, 
                                                                              fill = "#fbfbfb"))

@@

```{r ceofs-2, fig.cap = "Time series of the leading two cEOFs of zonal anomalies of geopotential height at 50 hPa and 200 hPa for the SON trimester and the period 1979 -- 2019.", fig.height=width_column, fig.width = width_column}
ceof %>%
  .[, eof[[1]]$left, by = season] %>%
  sep_ReIm(hgt) %>%
  .[season == "SON"] %>%
  # .[, hgt := scale(hgt), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt)) +
  geom_line(aes(color = part)) +
  scale_x_datetime(NULL) +
  scale_y_continuous(NULL) +
  scale_color_brewer(NULL, palette = "Dark2") +
  facet_grid(cEOF ~ ., switch = "y") +
  theme(panel.spacing = grid::unit(1.5, "lines"), strip.placement = "outside")  +

  panel_background
```

@@

(ref:extended-series-cap) Time series extended using ERA5 back extended preliminary edition (period 1950 -- 1978) and ERA5 (period 1979 -- 2019). Dark straight line is the OLS trend. Black horizontal and vertical line mark the mean value and range of each time series respectively.

```{r extended-series, fig.cap = "(ref:extended-series-cap)", fig.height=4, fig.width = 7, fig.env = "figure*"}
# Check that the extended index coincides with the
# original index.
cors <- era_more %>%
  .[sep_ReIm(ceof$eof[[1]]$left), on = c("time", "cEOF", "part")] %>%
  na.omit() %>%
  .[, cor(hgt, i.hgt), by = .(cEOF, part)]

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
   # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "r", size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_datetime(NULL, date_labels = "%Y") +
  scale_y_continuous("EOF", minor_breaks = NULL) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"))  +
  tag_facets("cr") +
  panel_background
```

@@

@@

The spatial patterns shown in Figure \@ref(fig:ceofs-1) are derived by removing the zonally symmetric circulation, so they might not include all the variability that is actually associated with the cEOF time series.
To understand the full geopotential height field changes associated with each cEOF, we computed regression patterns of the full geopotential fields onto the 
of each cEOF with the complete geopotential fields.

@@

  scale_x_level(minor_breaks = NULL) +
  scale_y_continuous(r2, limits = c(0, 1)) +
  scale_size_manual("p-value", values = c("TRUE" = 1, 
                                      "FALSE" = 0.5), 
                    guide = "none") +
  scale_color_brewer(NULL, palette = "Dark2", labels = c(full = "SAM",
                                                         asym = "A-SAM",
                                                         sym = "S-SAM")) +
  facet_grid(part ~ cEOF) +
  coord_flip() +
  tag_facets("rc") +
  panel_background
```
